---
title: "In class demos/exercises"
author: "Colin Rundel"
date: "2/01/2017"
output: html_document
---

```{r}
library(lubridate)
library(dplyr)
library(nycflights13)
library(ggplot2)
```

## Demos

### Demo 1

How many flights to Los Angeles (LAX) did each of the legacy carriers (AA, UA, DL or US) have in May from JFK, and what was their average duration?

```{r}
flights %>%
  filter(dest == "LAX") %>% 
  filter(carrier == "AA" | carrier == "UA" | carrier == "DL" | carrier == "US") %>%
  filter(month == 5) %>% 
  filter(origin == "JFK") %>%
  filter(!is.na(air_time)) %>%
  group_by(carrier) %>%
  summarize(n_flight = n(), mean_duration = mean(air_time))


#flights %>% filter(origin=="JFK",dest=="LAX",carrier=="US")
```

```{r}
flights %>%
  filter(carrier %in% c("AA", "UA", "DL", "US")) %>% 
  mutate(date = mdy(paste(month,day,year,sep="/"))) 
```

### Demo 2

Create a time series plot of each of the legacy carriers' average departure delay by day of the week and origin airport.

```{r}
dd = flights %>%
  filter(carrier %in% c("AA", "UA", "DL", "US")) %>% 
  mutate(date = mdy(paste(month,day,year,sep="/"))) %>%
  mutate(wday = wday(date, label=TRUE)) %>%
  group_by(carrier, wday, origin) %>% 
  summarize(avg_depdelay = mean(dep_delay, na.rm=TRUE))

ggplot(dd, aes(x=wday, y=avg_depdelay)) + 
  geom_point() +
  facet_grid(origin~carrier)
```



## Exercises

### Exercise 1

Which plane (check the tail number) flew out of each New York airport the most?

```{r}
flights %>%
  filter(!is.na(tailnum)) %>%
  group_by(tailnum, origin) %>%
  summarize(n=n()) %>%
  arrange(origin, desc(n), tailnum) %>%
  group_by(origin) %>%
  slice(1)
```

### Exercise 2

What was the shortest flight out of each airport in terms of distance? In terms of duration?

```{r}
flights %>% 
  select(origin, dest, flight, air_time, distance) %>%
  arrange(origin, air_time) %>%
  group_by(origin) %>% 
  slice(1)

flights %>% 
  select(origin, dest, flight, air_time, distance) %>%
  arrange(origin, distance) %>%
  group_by(origin) %>% 
  slice(1)
```

### Exercise 3

Which date should you fly on if you want to have the lowest possible average departure delay? What about arrival delay?

```{r}
flights %>%
  group_by(year, month, day) %>% 
  summarize(avg_depdelay = mean(dep_delay, na.rm=TRUE)) %>%
  arrange(avg_depdelay) %>%
  ungroup() %>%
  slice(1)

flights %>%
  group_by(year, month, day) %>% 
  summarize(avg_arrdelay = mean(arr_delay, na.rm=TRUE)) %>%
  arrange(avg_arrdelay) %>%
  ungroup() %>%
  slice(1)
```
